On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-06-13 main.tex
minor typos

\begin{document}

@@

Put $V = \bm\Phi^{-1}(\bm 0)$. If $V$ is empty, there is nothing to do,
since all values $\thetab$ in $\C^\dt$ satisfy the conclusion of the
proposition. Thus, we assume that $V$ is not empty. Then, the Jacobian
criterion shows that $V$ is smooth and $(n+ \dt -m)$-equidimensional.

@@

To describe the set $Z$ of critical points of $\pi_{|V}$, let $\bm M$
denote the $(s+m) \times (s+n)$ Jacobian matrix with entries in
$\C[\Xb,\Thetab]$ given by $\bm M = \jac_{\Xb,\Thetab}(\pi,\bm\Phi)$,
that is,
\begin{align*}
\bm M &= 
\bbm 
\jac_{\Xb,\Thetab}(\pi)\\
\jac_{\Xb,\Thetab}(\bm\Phi) 
\ebm 
=
\bbm 
\textbf{0}_{\dt \times n}\hspace{5mm}\textbf{I}_{\dt} \\
\jac_{\Xb,\Thetab}(\bm\Phi)
\ebm.
\end{align*}

@@

For this, we let $\textit{\textbf{L}}=[L_1,\hdots,L_m]$ be new
variables, thought of as Lagrange multipliers, and consider the
``Lagrange polynomials'' $\mathscr{L}_1,\dots,\mathscr{L}_{n}$, with
\[
[\mathscr{L}_1 \cdots \mathscr{L}_{n}]=\textit{\textbf{L}}\cdot \bm J(\xb,\thetab).
\] 
Denote by $\frak{Z} \subset \C^{n+\dt+m}$ the algebraic set defined by
the vanishing of $\mathscr{L}_1,\hdots,\mathscr{L}_{n},$ and
$\bm\Phi$, and by $\frak{Z}'$ the algebraic set
\[
\frak{Z}' := \overline{\frak{Z} - \{(\xb,\thetab,0,\hdots,0) \in \C^{n+\dt+m}~|~(\xb,\thetab,0\hdots,0) \in \frak{Z}\}},
\]
where the bar denotes Zariski closure (we have to remove such points,
since $L_1=\cdots=L_m=0$ is always a trivial solution to the Lagrange
equations). Finally, consider the projection
\begin{align*} 
\mu :~ \C^{n+\dt+m} &\rightarrow \C^{n+\dt}\\
(\xb,\thetab,\bm \ell)~ &\mapsto (\xb,\thetab).
\end{align*}

@@

In Section~\ref{sec:Hi2}, we will need the following by-product of
this result: if we consider $f^\A \in \C(\frak
A_{j,k})[X_1,\dots,X_n]$ as defined Section~\ref{sec:gen}, this
polynomial satisfies the rank property $\textbf{H}_i(1)$.

@@

\begin{lemma} 
  For $j=1,\dots,n$, there exists a polynomial $P_j$ in
  $\C(\A)[X_1,\dots,X_{i-1},X_j]$, monic in $X_j$, with $\pjb$ in
  $\I^\A$, and such that $\deg(\pjb)\leq (2d)^{n}.$
\end{lemma} 
\begin{proof}
  We let $\frak L^\A$ denote the extension of $\I^\A$ given by $\frak
  L^\A:=\I^\A\cdot \C(\frak A, X_1,\dots,X_{i-1})[X_i,\dots,X_n]$. Then,
  \begin{equation}\label{eq:1}
    \C(\frak A, X_1,\dots,X_{i-1}) \to \C(\frak A,
    X_1,\dots,X_{i-1})[X_i,\dots,X_n]/\frak L^\A
  \end{equation}
  is an algebraic extension. On the other hand, 
  the previous lemma states that 
  \begin{equation}\label{eq:2}
    \C(\A)[X_1,\dots,X_{i-1}]\rightarrow
    \C(\A)[X_1,\dots,X_n]/\I^{\A}  
  \end{equation}
  is integral; from this, Proposition~3.3.1 in~\cite{GiHeSa93} implies
  that it is actually a free module. Any basis of the latter is also a
  basis of~\eqref{eq:1}; as a consequence, for $j$ in $i,\dots,n$, the
  characteristic polynomials of $X_j$ in~\eqref{eq:1} or~\eqref{eq:2}
  are the same. Let $P_j$ be the minimal polynomial of $X_j$
  in~\eqref{eq:1}. The previous discussion implies that the
  characteristic polynomial $\chi_j$ of $X_j$ in~\eqref{eq:1}, and
  thus also $P_j$, are in $\C(\A)[X_1,\dots,X_{i-1},X_j]$ and monic in
  $X_j$.

@@

Our next step is to give degree bounds on the coefficients appearing
in the membership equality $\pjb \in\I^\A$. This is done using
Rabinovicz's trick. Let $T$ be a new variable; applying the
Nullstellensatz in $\C(\A)[X_1,\dots,X_n,T]$, and clearing
denominators, we obtain the existence of $\alpha_j$ in
$\C[\A]-\{0\}$ and $ C_{j,\ell},B_j$ in
$\C[\A][X_1,\dots,X_n][T]$, such that
\begin{align}\label{eq:nullst}
\alpha_j = \sum_{\ell=1}^{n-i+1} C_{j,\ell} G_\ell + B_j (1-\pjb T),\ \  G_\ell \in 
\left\{ 
f^\A,\frac{\partial f^\A}{\partial X_{i+1}},\hdots,\frac{\partial f^\A}{\partial X_n}
\right\}.
\end{align}

Let us then define 
$$\D_{i}:=\D_{i,1} \alpha_i \cdots \alpha_n D_i \cdots D_n,$$ where
$\D_{i,1}$ was defined in Section~\ref{ssec:Hi1} and for all $j$,
$\alpha_j$ is as above and $D_j$ is the leading coefficient of $\pjb$
with respect to $X_j$. Thus, $\D_{i}$ is a non-zero polynomial in
$\C[\A]$; we will estimate its degree below.

>>>>

\ccsdesc[500]{Computing methodologies~Algebraic algorithms}
\begin{document}

@@

Put $V := \bm\Phi^{-1}(\bm 0)$. If $V$ is empty, there is nothing to do,
since all values $\thetab$ in $\C^\dt$ satisfy the conclusion of the
proposition. Thus, we assume that $V$ is not empty. Then, the Jacobian
criterion shows that $V$ is smooth and $(n+ \dt -m)$-equidimensional.

@@

To describe the set $Z$ of critical points of $\pi_{|V}$, let $\bm M$
denote the $(s+m) \times (s+n)$ Jacobian matrix with entries in
$\C[\Xb,\Thetab]$ given by $\bm M := \jac_{\Xb,\Thetab}(\pi,\bm\Phi)$,
that is,
\begin{align*}
\bm M &= 
\bbm 
\jac_{\Xb,\Thetab}(\pi)\\
\jac_{\Xb,\Thetab}(\bm\Phi) 
\ebm 
=
\bbm 
\textbf{0}_{\dt \times n}\hspace{5mm}\textbf{I}_{\dt} \\
\jac_{\Xb,\Thetab}(\bm\Phi)
\ebm.
\end{align*}

@@

For this, we let $\textit{\textbf{L}}=[L_1,\hdots,L_m]$ be new
variables, thought of as Lagrange multipliers, and consider the
``Lagrange polynomials'' $\mathscr{L}_1,\dots,\mathscr{L}_{n}$, with
\[
[\mathscr{L}_1 \cdots \mathscr{L}_{n}]:=\textit{\textbf{L}}\cdot \bm J(\xb,\thetab).
\] 
Denote by $\frak{Z} \subset \C^{n+\dt+m}$ the algebraic set defined by
the vanishing of $\mathscr{L}_1,\hdots,\mathscr{L}_{n},$ and
$\bm\Phi$, and by $\frak{Z}'$ the algebraic set
\[
\frak{Z}' := \overline{\frak{Z} - \{(\xb,\thetab,0,\hdots,0) \in \C^{n+\dt+m}~|~(\xb,\thetab,0\hdots,0) \in \frak{Z}\}},
\]
where the bar denotes Zariski closure (we have to remove such points,
since $L_1=\cdots=L_m=0$ is always a trivial solution to the Lagrange
equations). Finally, consider the projection
\begin{align*} 
\mu :~ \C^{n+\dt+m} &\rightarrow \C^{n+\dt}\\
(\xb,\thetab,\bm \ell)~ &\mapsto (\xb,\thetab).
\end{align*}

@@

In Section~\ref{sec:Hi2}, we will need the following by-product of
this result: if we consider $f^\A \in \C(\frak
A_{j,k})[X_1,\dots,X_n]$ as defined in Section~\ref{sec:gen}, this
polynomial satisfies the rank property $\textbf{H}_i(1)$.

@@

\begin{lemma} 
  For $j=i,\dots,n$, there exists a polynomial $P_j$ in
  $\C(\A)[X_1,\dots,X_{i-1},X_j]$, monic in $X_j$, with $\pjb$ in
  $\I^\A$, and such that $\deg(\pjb)\leq (2d)^{n}.$
\end{lemma} 
\begin{proof}
  We let $\frak L^\A$ denote the extension of $\I^\A$ given by $\frak
  L^\A:=\I^\A\cdot \C(\frak A, X_1,\dots,X_{i-1})[X_i,\dots,X_n]$. Then,
  \begin{equation}\label{eq:1}
    \C(\frak A, X_1,\dots,X_{i-1}) \to \C(\frak A,
    X_1,\dots,X_{i-1})[X_i,\dots,X_n]/\frak L^\A
  \end{equation}
  is an algebraic extension. On the other hand, 
  the previous lemma states that 
  \begin{equation}\label{eq:2}
    \C(\A)[X_1,\dots,X_{i-1}]\rightarrow
    \C(\A)[X_1,\dots,X_n]/\I^{\A}  
  \end{equation}
  is integral; from this, Proposition~3.3.1 in~\cite{GiHeSa93} implies
  that it is actually a free module. Any basis of the latter is also a
  basis of~\eqref{eq:1}; as a consequence, for $j$ in $i,\dots,n$, the
  characteristic polynomials of $X_j$ in~\eqref{eq:1} or~\eqref{eq:2}
  are the same. Let $P_j$ be the minimal polynomial of $X_j$
  in~\eqref{eq:1}. The previous discussion implies that the
  characteristic polynomial $\chi_j$ of $X_j$ in~\eqref{eq:1}, and
  thus also $P_j$, are in $\C(\A)[X_1,\dots,X_{i-1},X_j]$ and monic in
  $X_j$.

@@

Our next step is to give degree bounds on the coefficients appearing
in the membership equality $\pjb \in\I^\A$. This is done using
Rabinovicz's trick. Let $T$ be a new variable; applying the
Nullstellensatz in $\C(\A)[X_1,\dots,X_n,T]$, and clearing
denominators, we obtain the existence of $\alpha_j$ in
$\C[\A]-\{0\}$ and $ C_{j,\ell},B_j$ in
$\C[\A][X_1,\dots,X_n][T]$, such that
%
\begin{align}\label{eq:nullst}
\alpha_j = \sum_{\ell=1}^{n-i+1} C_{j,\ell} G_\ell + B_j (1-\pjb T),\ \  G_\ell \in 
\left\{ 
f^\A,\frac{\partial f^\A}{\partial X_{i+1}},\hdots,\frac{\partial f^\A}{\partial X_n}
\right\}.
\end{align}
%
Let us then define 
$$\D_{i}:=\D_{i,1} \alpha_i \cdots \alpha_n D_i \cdots D_n,$$ where
$\D_{i,1}$ was defined in Section~\ref{ssec:Hi1} and for all $j$,
$\alpha_j$ is as above and $D_j$ is the leading coefficient of $\pjb$
with respect to $X_j$. Thus, $\D_{i}$ is a non-zero polynomial in
$\C[\A]$; we will estimate its degree below.
